---
title: "explore CPAP adherence"
author: "Xing Song"
date: "03/01/2022"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE,
                      warning = FALSE,
                      fig.height=5)

pacman::p_load(tidyverse,
               magrittr,
               broom,
               survival,
               survminer,
               kableExtra,
               plotly,
               # threejs,
               # rgl,
               ggpubr)
source("./util.R")
```


```{r, load_data}
dat<-readRDS("./data/analytic_set.rda") %>%
  mutate(DAYS_CPAP_TO_SUPPLY = DAYS_ENR_START_TO_SUPPLY - DAYS_ENR_START_TO_CPAP,
         MTHS_CPAP_TO_SUPPLY = round(DAYS_CPAP_TO_SUPPLY/30))
# take out unique patient list
pat_uni<-dat %>% select(PATID_DEID) %>% unique
```

Let's first inspect some timeline examples: 

```{r, plot_sample}
sample<-dat %>%
  semi_join(pat_uni %>% sample_n(10),by="PATID_DEID")

ggplot(data=sample, aes(group = PATID_DEID)) +
  geom_line(aes(x=DAYS_ENR_START_TO_SUPPLY,y=as.factor(PATID_DEID)))+
  geom_point(aes(x=DAYS_ENR_START_TO_SUPPLY,y=as.factor(PATID_DEID)))+
  geom_point(aes(x=DAYS_ENR_START_TO_OSA,y=as.factor(PATID_DEID)),color='red')+
  geom_point(aes(x=DAYS_ENR_START_TO_CPAP,y=as.factor(PATID_DEID)),color='green')+
  geom_point(aes(x=DAYS_ENR_START_TO_END,y=as.factor(PATID_DEID)),color='blue')+
  labs(x="Days since enrollment start",y="Random Patient ID")
```

Let's then switch view to an Area Under the Curve Measurement of cumulative CPAP charge vs. Days from CPAP initiation (so that time origins are aligned at CPAP initiation): 

```{r, sample_auc}
ggplot(data=sample, aes(x=DAYS_CPAP_TO_SUPPLY, y = CPAP_CUM_DOSE)) +
  geom_line()+ geom_point() +
  geom_vline(xintercept = 365, color = "red", linetype = 2) +
  geom_vline(xintercept = 365*2, color = "red", linetype = 2) +
  geom_vline(xintercept = 365*3, color = "red", linetype = 2) +
  facet_wrap(~ PATID_DEID, ncol = 5, scale = "free") +
  labs(x="Days since enrollment start",y="Cumulative CPAP charges (number of claims)")

```

```{r, surface_plot, webgl=TRUE}
cpap_freq<-dat %>%
  group_by(CPAP_CUM_DOSE, MTHS_CPAP_TO_SUPPLY) %>%
  summarize(freq = length(unique(PATID_DEID)),.groups = "drop") %>%
  filter(freq >= 5)

p<-plot_ly() %>%
  add_trace(data = cpap_freq,
            x = cpap_freq$CPAP_CUM_DOSE,
            y = cpap_freq$MTHS_CPAP_TO_SUPPLY,
            z = cpap_freq$freq, type = "mesh3d") %>%
  layout(scene = list(xaxis = list(title = "Cumulative CPAP charges"),
                      yaxis = list(title = "Months since CPAP initiation"),
                      zaxis = list(title = "Patient frequencies")))
htmlwidgets::saveWidget(as_widget(p), "surface_plot.html")

# plot3d(cpap_freq$CPAP_CUM_DOSE,cpap_freq$MTHS_CPAP_TO_SUPPLY,cpap_freq$freq,
#           col=rainbow(length(cpap_freq$freq)))

```

Based on both the sample and the surface plot, it is obvious that the usage pattern is often a lot more intense in CPAP initiation year compared to subsequent years. However, not all patients get to be followed up for more than a year before censored (e.g., last data available year in 2017).   

```{r}
N<-pat_uni %>% nrow()
n1<-pat_uni %>% semi_join(dat %>% filter(DAYS_CPAP_TO_SUPPLY>365),by="PATID_DEID") %>% nrow()
p1<-round(n1/N,2)
n2<-pat_uni %>% semi_join(dat %>% filter(DAYS_CPAP_TO_SUPPLY>365*2),by="PATID_DEID") %>% nrow()
p2<-round(n2/N,2)
gc()
```

Among `r N` eligible patients, about `r n1`(`r p1`) patients is followed for at least 1 complete year post CPAP initiation, about `r n2`(`r p2`) is followed up for at least 2 complete years post CPAP initiation. Following Wickwire's definition, i.e., total number of CPAP charges, the following figure shows what the frequency distribution looks like: 

```{r}
dat2<-readRDS("./data/analytic_set2.rda") 

p1<-ggplot(dat2,aes(x = adherence_ww)) + 
  geom_bar(aes(y =  (..count..)/sum(..count..), text = ..count..),fill = "red") + 
  scale_y_continuous(labels=scales::percent)

p2<-ggplot(dat2 %>% semi_join(dat %>% filter(DAYS_CPAP_TO_SUPPLY>365),by="PATID_DEID"),
           aes(x = adherence_ww)) + 
  geom_bar(aes(y =  (..count..)/sum(..count..), text = ..count..), fill = "blue") + 
  scale_y_continuous(labels=scales::percent)

ggarrange(p1, p2, ncol = 2, nrow = 1)

```

A significant amount of "low" adherence is removed as a result of non-sufficient follow-up period (less than 1 year). For those been followed up for more than 1 year, more than 90% are high-compliance based on Wickwire's definition. It might be more interesting to further breakdown the "high adherence" group into finer groups using empirical metrics such as percentil rank. 


